---
title: "Spotify_YS"
output:
  html_document: default
  word_document: default
  pdf_document: default
date: "2023-04-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Importing the the useful packages
```{r , include=TRUE , warning=TRUE}
library(readr)
library(tidyverse)
library(corrplot)
library(ggplot2)
library(pROC)
library(rpart)
library(tree)
```

```{r }
library(rpart.plot)
library(caret)
```
# Importing data


```{r }
data_YT <- read_csv("C:/Users/hp/Desktop/freelancer work/Spotify_Youtube/data.csv")
```
# Manipulating data
```{r }
head(data_YT)
```
## Missing Value

```{r }
MissingValue<-colSums(is.na(data_YT))
MissingValue
```

##  Removing Missing Value
```{r }
data_YT<-na.omit(data_YT)
```
# Statistics descriptive and data exploratory
```{r }
statistics<-data_YT%>%
  sapply(summary)%>%
  as.data.frame()
statistics<-t(statistics)%>%
  as.data.frame()
statistics[,"sd"]<-data_YT%>%
  sapply(sd)%>%
  as.data.frame()
print(statistics)

```
## correlation

```{r }
corrplot(cor(data_YT),method="square",addCoef.col="black",type="upper", tl.cex=.8, number.cex=.5)
```

## Histograms

#### danceability
```{r }

data_YT$Song_Liked<-ifelse(data_YT$liked==1,"Liked","Disliked")

ggplot(data_YT, aes(x=danceability,fill=Song_Liked,col=Song_Liked))+
  geom_histogram( position="identity",bins=20)+
  theme_minimal()+
  labs(x='danceability', y='Frequency', title='Histogram of The danceability by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```

```{r }
ggplot(data_YT, aes(energy, fill=Song_Liked)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")

```
#### Energy
```{r }
ggplot(data_YT, aes(x=energy,fill=Song_Liked,col=Song_Liked))+
  geom_histogram( position="identity",bins=20)+
  theme_minimal()+
  labs(x='Energy', y='Frequency', title='Histogram of The Energy by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```

```{r }
ggplot(data_YT, aes(energy, fill=Song_Liked)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")

```

#### Key
```{r }
ggplot(data_YT, aes(x=key,fill=Song_Liked,col=Song_Liked))+
  geom_bar( position="identity")+
  theme_minimal()+
  labs(x='key', y='Frequency', title='Histogram of The kety by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```

```{r }
ggplot(data_YT, aes(key, fill=Song_Liked)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")


```
#### Loudness
```{r }


ggplot(data_YT, aes(x=loudness,fill=Song_Liked,col=Song_Liked))+
  geom_histogram( position="identity",bins=20)+
  theme_minimal()+
  labs(x='loudness', y='Frequency', title='Histogram of The Loudness by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```

```{r }
ggplot(data_YT, aes(energy, fill=Song_Liked)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")

```
#### Speechiness
```{r }
ggplot(data_YT, aes(x=speechiness,fill=Song_Liked,col=Song_Liked))+
  geom_histogram( position="identity",bins=20)+
  theme_minimal()+
  labs(x='Speechiness', y='Frequency', title='Histogram of The Speechiness by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```
```{r }
ggplot(data_YT, aes(speechiness, fill=Song_Liked)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")

```

#### Acousticness
```{r }
ggplot(data_YT, aes(x=acousticness,fill=Song_Liked,col=Song_Liked))+
  geom_histogram( position="identity",bins=20)+
  theme_minimal()+
  labs(x='Acousticness', y='Frequency', title='Histogram of The Acousticness by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```

```{r }
ggplot(data_YT, aes(acousticness, fill=Song_Liked)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")

```
#### Instrumentalness
```{r }


ggplot(data_YT, aes(x=instrumentalness,fill=Song_Liked,col=Song_Liked))+
  geom_histogram( position="identity",bins=20)+
  theme_minimal()+
  labs(x='Instrumentalness', y='Frequency', title='Histogram of The Instrumentalness by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```

```{r }
ggplot(data_YT, aes(instrumentalness, fill=Song_Liked)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")

```

#### Liveness
```{r }
ggplot(data_YT, aes(x=liveness,fill=Song_Liked,col=Song_Liked))+
  geom_histogram( position="identity",bins=20)+
  theme_minimal()+
  labs(x='Liveness', y='Frequency', title='Histogram of The Liveness by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```

```{r }
ggplot(data_YT, aes(liveness, fill=Song_Liked)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")
```

#### Valence
```{r }
ggplot(data_YT, aes(x=valence,fill=Song_Liked,col=Song_Liked))+
  geom_histogram( position="identity",bins=20)+
  theme_minimal()+
  labs(x='Valence', y='Frequency', title='Histogram of The Valence by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```

```{r }
ggplot(data_YT, aes(valence, fill=Song_Liked)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")
```

#### tempo
```{r }


ggplot(data_YT, aes(x=tempo,fill=Song_Liked,col=Song_Liked))+
  geom_histogram( position="identity",bins=20)+
  theme_minimal()+
  labs(x='Tempo', y='Frequency', title='Histogram of The Tempo by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```

```{r }
ggplot(data_YT, aes(tempo, fill=Song_Liked)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")

```

#### Duration_ms
```{r }
ggplot(data_YT, aes(x=duration_ms,fill=Song_Liked,col=Song_Liked))+
  geom_histogram( position="identity",bins=20)+
  theme_minimal()+
  labs(x='Duration_ms', y='Frequency', title='Histogram of The Duration_ms by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```

```{r }
ggplot(data_YT, aes(duration_ms, fill=Song_Liked)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")

```

#### time_signature
```{r }
ggplot(data_YT, aes(x=time_signature,fill=Song_Liked,col=Song_Liked))+
  geom_bar( position="identity")+
  theme_minimal()+
  labs(x='time_signature', y='Frequency', title='Histogram of The time_signature by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```

```{r }
ggplot(data_YT, aes(time_signature, fill=Song_Liked)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")

```
#### Time_signature
```{r }
ggplot(data_YT, aes(x=liked,fill=Song_Liked,col=Song_Liked))+
  geom_bar(stat="bin",binwidth=0.65)+
  theme_minimal()+
  labs(y='Frequency', title='Liked vs Disliked songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```

```{r }
ggplot(data_YT, aes(time_signature, fill=Song_Liked)) + 
  geom_density(alpha=.5) + 
  scale_fill_manual(values = c('#999999','#E69F00')) + 
  theme(legend.position = "none")


```
## Boxplot

##### danceability
```{r }
ggplot(data_YT, aes(y=danceability, x=Song_Liked,fill=Song_Liked))+
  geom_boxplot( position="identity")+
  theme_minimal()+
  labs( y='Danceability', title='Box plot of The danceability by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```
#### Loudness
```{r }
ggplot(data_YT, aes(y=loudness,fill=Song_Liked,x=Song_Liked))+
  geom_boxplot( position="identity")+
  theme_minimal()+
  labs( y='Loudness', title='Box plot of The Loudness by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```

#### Speechiness
```{r }
ggplot(data_YT, aes(y=speechiness,fill=Song_Liked,x=Song_Liked))+
  geom_boxplot( position="identity")+
  theme_minimal()+
  labs( y='Speechiness', title='Box plot of The Speechiness by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```
#### Instrumentalness
```{r }

ggplot(data_YT, aes(y=instrumentalness,fill=Song_Liked,x=Song_Liked))+
  geom_boxplot( position="identity")+
  theme_minimal()+
  labs( y='instrumentalness', title='Box plot of The instrumentalness by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```
#####Duration_ms
```{r }
ggplot(data_YT, aes(y=duration_ms,fill=Song_Liked,x=Song_Liked))+
  geom_boxplot( position="identity")+
  theme_minimal()+
  labs( y='Duration_ms', title='Box plot of The Duration by Liked & Disliked Songs')+
  theme(plot.title = element_text(hjust=0.5, size=10, face='bold'),legend.position = "bottom")

```
# Split the data
```{r }
data_YT <- read_csv("C:/Users/hp/Desktop/freelancer work/Spotify_Youtube/data.csv")
n=dim(data_YT)[1]
set.seed(1251)
train.index=sample(n, round(0.7*n))
train=data_YT[train.index, ]
test=data_YT[-train.index, ]

```
# Machine learning algorithms
## Logistic regression

#### Model 0
```{r }
logit.model=glm(liked~.,data=train,family="binomial")
summary(logit.model)

```

```{r }
prob_logit<- predict(logit.model,newdata=test,type="response")
logit.pred <- ifelse(prob_logit > 0.5, 1, 0)
logit.sum=confusionMatrix(data=as.factor(logit.pred), reference= as.factor(test$liked), positive="1")

```

```{r }
logit.sum

```

```{r }
roc(logit.pred,test$liked)
plot.roc(logit.pred, test$liked)

```

#### Model 1
```{r }

logit.model1=glm(liked~danceability +energy+key+loudness+mode+speechiness+liveness+valence+tempo+duration_ms+time_signature,data=train,family="binomial")
summary(logit.model1)

```

```{r }
prob_logit1<- predict(logit.model1,newdata=test,type="response")
logit1.pred <- ifelse(prob_logit1 > 0.5, 1, 0)
logit1.sum=confusionMatrix(data=as.factor(logit1.pred), reference= as.factor(test$liked), positive="1")
logit1.sum

```

```{r }
roc(logit1.pred,test$liked)

```

```{r }
plot.roc(logit1.pred, test$liked)


```

#### Model 2
```{r }

logit.model2=glm(liked~danceability +energy+loudness+tempo+duration_ms,data=train,family="binomial")

summary(logit.model2)

```

```{r }
glm.predict=predict(logit.model2,newdata=test,type="response")

prob_logit2<- predict(logit.model2, test,type = "response")
logit2.pred <- ifelse(prob_logit2 > 0.5, 1, 0)
logit2.sum=confusionMatrix(data=as.factor(logit2.pred), reference= as.factor(test$liked), positive="1")
logit2.sum

```
```{r }
roc(logit2.pred,test$liked)

```

```{r }
plot.roc(logit2.pred, test$liked)

```

## linear discriminant analysis
#### Model 0
```{r }
library(MASS)
lda0.fit = lda(liked ~ ., data =train)
lda0.pred = predict(lda0.fit, test)
lda0.sum=confusionMatrix(data=as.factor(lda0.pred[["class"]]), reference= as.factor(test$liked), positive="1")
lda0.sum
```

```{r }
roc(lda0.pred[["class"]], test$liked)
```

```{r }
plot.roc(lda0.pred[["class"]], test$liked)
```
#### Model 1
```{r }

lda1.fit = lda(liked ~ danceability +energy+loudness+tempo+duration_ms, data =train)
lda1.pred = predict(lda1.fit, test)
lda1.sum=confusionMatrix(data=as.factor(lda1.pred[["class"]]), reference= as.factor(test$liked), positive="1")
lda1.sum
```

```{r }
roc(lda1.pred[["class"]], test$liked)
```

```{r }
plot.roc(lda1.pred[["class"]], test$liked)
```
## quadratic discriminant analysis
#### Model 0
```{r }
qda0.fit = qda(liked ~ ., data = train)
qda0.class = predict(qda0.fit, test)$class
qda0.sum=confusionMatrix(data=as.factor(qda0.class), reference= as.factor(test$liked), positive="1")
qda0.sum
```

```{r }
roc(qda0.class, test$liked)
```

```{r }
plot.roc(qda0.class, test$liked)
```
#### Model 1
```{r }

qda1.fit = qda(liked ~danceability +energy+loudness+tempo+duration_ms, data = train)
qda1.class = predict(qda1.fit, test)$class
qda1.sum=confusionMatrix(data=as.factor(qda1.class), reference= as.factor(test$liked), positive="1")
qda1.sum
```

```{r }
roc(qda1.class, test$liked)
```

```{r }
plot.roc(qda1.class, test$liked)
```
## K nearest neighbor
#### Model 0
```{r }
train.X = train[, -14]
test.X = test[, -14]
train.y = train$liked

k1=round((sqrt(NROW(train))),0)

library(class)
knn0.class = knn(train.X, test.X, train.y, k = 11)
knn0.sum=confusionMatrix(data=as.factor(knn0.class), reference= as.factor(test$liked), positive="1")
knn0.sum
```

```{r }
roc(knn0.class, test$liked)
```

```{r }
plot.roc(knn0.class, test$liked)
```
#### Model 1
```{r }


knn1.class = knn(train.X, test.X, train.y, k = 12)
knn1.sum=confusionMatrix(data=as.factor(knn1.class), reference= as.factor(test$liked), positive="1")
knn1.sum
```

```{r }
roc(knn1.class, test$liked)
```

```{r }
plot.roc(knn1.class, test$liked)
```
## Decision Tree
#### Model 0
```{r }


#tree0
tree <- rpart(liked ~., data = train)
rpart.plot(tree)
```

```{r }
printcp(tree)
```

```{r }
plotcp(tree)
```

```{r }
# Make predictions
probabilities <- predict(tree, test.X)
predicted.classes <- ifelse(probabilities > 0.5, 1, 0)
# confusionMatrix
tree.sum=confusionMatrix(data=as.factor(predicted.classes), reference= as.factor(test$liked), positive="1")
tree.sum
```

```{r }
roc(predicted.classes, test$liked)
```

```{r }
plot.roc(predicted.classes, test$liked)
```
#### Model 1
```{r }

#tree
tree1 <- rpart(liked ~., data = train,cp=0.034590)
rpart.plot(tree1)
```

```{r }
# Make predictions
probabilities1 <- predict(tree1, test.X)
predicted.classes1 <- ifelse(probabilities1 > 0.5, 1, 0)
# confusionMatrix
tree1.sum=confusionMatrix(data=as.factor(predicted.classes1), reference= as.factor(test$liked), positive="1")
tree1.sum
```

```{r }
roc(predicted.classes1, test$liked)
```

```{r }
plot.roc(predicted.classes1, test$liked)
```

## Naive Bayes
#### Model 0
```{r }

library(e1071)

nb0.fit = naiveBayes(liked ~ ., data = train)
nb0.class = predict(nb0.fit, test)
nb0.sum=confusionMatrix(data=as.factor(nb0.class), reference= as.factor(test$liked), positive="1")
nb0.sum
```

```{r }
roc(nb0.class, test$liked)
```

```{r }
plot.roc(nb0.class, test$liked)
```
#### Model 0
```{r }



nb1.fit = naiveBayes(liked ~ danceability +energy+loudness+speechiness+tempo+duration_ms, data = train)
nb1.class = predict(nb1.fit, test)
nb1.sum=confusionMatrix(data=as.factor(nb1.class), reference= as.factor(test$liked), positive="1")
nb1.sum
```

```{r }
roc(nb1.class, test$liked)
```

```{r }
plot.roc(nb1.class, test$liked)
```
## Bagging
#### Model 0
```{r }
library(rpart)
library(caret)
library(ipred)
#bag0
bag <- bagging(liked ~., data = train, nbagg=100,coob=TRUE,control=rpart.control(minsplit=2,cp=0))
# Make predictions
prob_bag <- predict(bag, test.X)
pred.bag.classes <- ifelse(prob_bag > 0.5, 1, 0)
# confusionMatrix
bag.sum=confusionMatrix(data=as.factor(pred.bag.classes), reference= as.factor(test$liked), positive="1")
bag.sum
```

```{r }
roc(pred.bag.classes, test$liked)
```

```{r }
plot.roc(pred.bag.classes, test$liked)
```
#### Model 1
```{r }



bag1 <- bagging(liked ~danceability +energy+loudness+tempo+duration_ms, data = train, nbag1g=100,coob=TRUE,control=rpart.control(minsplit=2,cp=0))
# Make predictions
prob_bag1 <- predict(bag1, test.X)
pred.bag1.classes <- ifelse(prob_bag1 > 0.5, 1, 0)
# confusionMatrix
bag1.sum=confusionMatrix(data=as.factor(pred.bag1.classes), reference= as.factor(test$liked), positive="1")
bag1.sum
```

```{r }
roc(pred.bag1.classes, test$liked)
```

```{r }
VI1<-data.frame(var=setdiff(names(train),"liked"),imp=varImp(bag))
VI_plot<-VI1[order(VI1$Overall,decreasing = TRUE),]
barplot(VI_plot$Overall,names.arg = row.names(VI_plot),horiz = TRUE,col='lightblue',xlab = "Varaible Impotance",las=2, cex.names = 0.5)
```

```{r }
plot.roc(pred.bag1.classes, test$liked)
```

## Random forest
#### Model 0
```{r }
library(randomForest)
n_features<-length(setdiff(names(train),"liked"))
rf<-randomForest(liked ~., data = train,mtry=floor(n_features/3),importance=TRUE)
rf
```

```{r }
rf$importance
```

```{r }
plot(rf$mse,type='l',col=2,lwd=2,xlab="ntree",ylab="OOB Error")
```

```{r }

# Make predictions
prob_rf0 <- predict(rf, test.X)
pred.rf0.classes <- ifelse(prob_rf0 > 0.5, 1, 0)
# confusionMatrix
rf0.sum=confusionMatrix(data=as.factor(pred.rf0.classes), reference= as.factor(test$liked), positive="1")
rf0.sum
```

```{r }
roc(pred.rf0.classes, test$liked)
```
#### Model 1
```{r }



rf1<-randomForest(liked ~., data = train,mtry=floor(sqrt(n_features/3)),importance=TRUE)
rf1
```

```{r }
rf1$importance
```

```{r }
plot(rf1$mse,type='l',col=2,lwd=2,xlab="ntree",ylab="OOB Error")
```

```{r }

# Make predictions
prob_rf1 <- predict(rf1, test.X)
pred.rf1.classes <- ifelse(prob_rf1 > 0.5, 1, 0)
# confusionMatrix
rf1.sum=confusionMatrix(data=as.factor(pred.rf1.classes), reference= as.factor(test$liked), positive="1")
rf1.sum
```

```{r }
roc(pred.rf1.classes, test$liked)
```
